Cluster Difference Imaging Photometric Survey. III. An Open Cluster With a 250\,pc Halo

### 2021-01-30 cluster_paper/ms.tex
update paper

% Context, aims, methods, results, conclusions.
\begin{abstract}
  Recent analyses of the Gaia data have reported the existence of
  diffuse stellar populations (``halos'') surrounding nearby open clusters.  The stars in these
  halos could have escaped from the cores of the cluster, or they
  could be remnants of a star formation
  complex that has since dispersed.  They could also be field
  stars---false positives originating from overly aggressive clustering algorithms.
  In this study, we focus on a single reported 250\,pc-long halo
  around the $\approx$120\,Myr open cluster NGC\,2516.  The classical tidal radius of
  this cluster is $\sim$XX\,pc.  Combining photometry from Gaia,
  rotation periods from TESS, and lithium measurements from
  the Gaia-ESO and GALAH surveys, we find that the halo of
  NGC\,2516 is real and coeval with the core. At fixed stellar
  mass, the rotation rate and
  lithium abundance are strongly anti-correlated. And finally, given Gaia-selected halo
  members, the contamination rate of field stars in the halo (XX\%) is
  larger than that in the core (YY\%).  This work expands the set of
  confirmed NGC\,2516 members by a factor of $\approx$2, and
  quantifies the degree to which Gaia-based analyses can 
  identify dispersed stellar populations against a background of field
  star contaminants.  We highlight the implications for spectroscopic
  survey targeting, open cluster dispersal, and planet searches around
  young stars.
\end{abstract}

@@

@@

Figure~\ref{fig:gaia6d_x_rotn} shows the result.  An important nuance
in interpreting this plot is that we show only stars with
$0.5<(Bp-Rp)_0<1.2$ for which our TESS pipeline successfully made
detrended light curves.  In other words---the base ``core'' and
``halo'' samples are the stars for which we could have plausibly
detected a rotation period, if the star were really a cluster member.
Faint stars with $Rp>16$ were beyond our light curve selection
limit, and crowded stars ({\it e.g.,} near saturated stars
in the cluster center) do not appear.
We selected the color limit above based on Figure~\ref{fig:rot},
since the density of period detections seems to decrease once 
$(Bp-Rp)_0 \gtrsim 1.25$, {\it i.e.}, for spectral types later
than $\approx$K4V.

Qualitatively, the fraction of kinematic members with TESS-detection rotation 
periods with increasing distance from the halo core---both in
position and in velocity space.
To quantify this, we computed the distances.

>>>>

% Context, aims, methods, results, conclusions.
\begin{abstract}
  Recent analyses of the Gaia data have reported the existence of
  diffuse stellar populations (``halos'') surrounding nearby open clusters.  The stars in these
  halos could have escaped from the cores of the cluster, or they
  could be remnants of a star formation
  complex that has since dispersed.  They could also be field
  stars---the false positives of overly aggressive clustering algorithms.
  In this study, we focus on a single reported 250\,pc-long halo
  around the $\approx$120\,Myr open cluster NGC\,2516.  The classical tidal radius of
  this cluster is $\sim$XX\,pc.  Combining photometry from Gaia,
  rotation periods from TESS, and lithium measurements from
  the Gaia-ESO and GALAH surveys, we find that the halo of
  NGC\,2516 is real and coeval with the core. At fixed stellar
  mass, the rotation rate and
  lithium abundance are strongly anti-correlated. And finally, given Gaia-selected halo
  members, the contamination rate of field stars in the halo (XX\%) is
  larger than that in the core (YY\%).  This work expands the set of
  confirmed NGC\,2516 members by a factor of $\approx$2, and
  quantifies the degree to which Gaia-based analyses can 
  identify dispersed stellar populations against a background of field
  star contaminants.  We highlight the implications for spectroscopic
  survey targeting, open cluster dispersal, and planet searches around
  young stars.
\end{abstract}

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.75\textwidth]{f8a.pdf}
		}

		\subfloat{
			\includegraphics[width=0.75\textwidth]{f8b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf Gaia-based components of NGC\,2516 in position and
    velocity space, cross-matched against the rotators, but now in
    physical units.}
    Top: relevant projections. Bottom: binned versus 3d position and
    2d velocity difference.
		\label{fig:physical_x_rotn}
	}
\end{figure*}

@@

Figure~\ref{fig:gaia6d_x_rotn} shows the result.  An important nuance
in interpreting this plot is that we show only stars with
$0.5<(Bp-Rp)_0<1.2$ for which our TESS pipeline successfully made
detrended light curves.  In other words---the base ``core'' and
``halo'' samples are the stars for which we could have plausibly
detected a rotation period.
Faint stars with $Rp>16$ were beyond our light curve selection
limit, and crowded stars ({\it e.g.,} near saturated stars
in the cluster center) do not appear.
We selected the color limit above based on Figure~\ref{fig:rot},
since the density of period detections seems to decrease once 
$(Bp-Rp)_0 \gtrsim 1.25$, {\it i.e.}, for spectral types later
than $\approx$K4V.

Qualitatively, the fraction of kinematic members with TESS-detected rotation 
periods seems like it may increase with increasing distance from the halo core---both in
position and in velocity space.
To quantify this, we computed the relelvant distances in physical coordinates.
This entailed first transforming from $(\alpha, \delta, \pi)$ to galactocentric ($X,Y,Z$) which we did 
assuming the AstroPy v4.0 coordinate standard (CITE).
We then converted the on-sky proper motions to physical units by dividing by the distance of each star.
We then computed the SEPARATIONS of all of these positions and
2d-velocities from the ``median core member'', which was selected by
taking \citetalias{cantatgaudin_gaia_2018} members with membership
probability exceeding 70\%.
Again, we only considered stars for which we could have plausibly
detected a rotation period --- the same color and observability limits
discussed above were also applied.

The results are shown in Figure~\ref{fig:physical_x_rotn}.
The halo extends to about 250\,pc in physical space, and out to at
least 10 of km/s in tangential velocity space.
\citet{meingast_2021} by comparison required a physically motivated cut
in tangential velocity space of 1.5 km per sec.
This probably improves purity, but clearly doesn't get everthing for
completeness.

% [361  82  43  30  16  18  14   8   6   2   4   0   0   0   0   0]
% [495 116  65  46  28  20  22  12  11   8   5   1   0   1   0   2]
% [384 109  45  27  15   4   0   0   0   0   0   0   0   0   0   0   0   0
%    0   0]
% [533 153  67  40  23  10   3   1   2   0   0   0   0   0   0   0   0   0
%    0   0]
